### 2019-06-15 R/analysis.Rmd
Add markdown document

---
title: "Royal Society paper"
subtitle: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r load_packages, include=FALSE}

library(tidyverse)
library(lubridate)
library(broom)
library(broman)
library(morris)
```

Words