### 2019-06-20 R/analysis.Rmd
Add varcomp to rmd, update lm and rarefy scripts

---
title: "Royal Society paper"
subtitle: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

# Overview

This is the primary analysis script for the analysis associated with the Royal Society
paper.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r load_packages, include=FALSE}

library(tidyverse)
library(lubridate)
library(broom)
library(broman)
library(morris)
library(varComp)
library(vegan)

```

```{r import_data}

all_data <- read.csv('../output/gab_all_rare_troph.csv', stringsAsFactors = FALSE)

```

# Variance Component Analysis

```{r sim_mat}

## Remove ASVs only present in 1 sample
asvs <- as.matrix(all_data[, grep("^asv", colnames(all_data))])
asvs[asvs>0] <- 1
all_data <- all_data[rowSums(asvs) > 1, ]
asv_mat <- as.matrix(all_data[, grep("^asv", colnames(all_data))])

dim(asvs) - dim(asv_mat)
rowSums(asvs)

# Community similarity
asv <- 3
taxon <- asv_mat[, asv]
#com_dis <- as.matrix(dist(as.matrix(scale(asv_mat[, -asv]))))
#com_sim <- 1/(1+com_dis)
com_sim <- 1 - as.matrix(vegdist(asv_mat[, -asv]))

env_variables <- all_data[, 1:8]

# Function
lowk <- all_data$Low_final_k

# Geographic similarity
xy <- all_data[, c('X', 'Y')]
geo_dis <- as.matrix(dist(as.matrix(scale(xy))))
geo_sim <- 1/(1+geo_dis)

# environmental dissimilarity
wetland <- all_data[, "Wetland"]
env <- all_data[, c('WFPS', 'N_percent', 'C_percent')]
env_dis <- as.matrix(dist(as.matrix(scale(env))))
env_sim <- 1/(1+env_dis)

```

```{r varcomp_diagnostics}

mantel(com_sim, geo_sim)
qplot(x = xy$X, y = xy$Y) + geom_jitter(aes(color = all_data$Wetland))

ggplot(mapping = aes(lowk)) + 
  geom_histogram(bins = 20) + 
  labs(x = "Low Final K", y = "Number of Samples")
ggplot(mapping = aes(taxon)) + 
  geom_histogram(bins = 20) + 
  labs(x = "Taxon Abundance", y = "Number of Samples")

all_data[all_data$Y > 30000, 'geocode'] <- 'A'
all_data[all_data$Y < 30000 & all_data$Y > 10000, 'geocode'] <- 'B'
all_data[all_data$Y < 10000, 'geocode'] <- 'C'
all_data$geocode <- factor(all_data$geocode)

ggplot(mapping = aes(c(com_sim))) + 
  geom_histogram(binwidth = 0.05) + 
  labs(x = "Community Similarity (Bray-Curtis)", y = "Number of Samples")
ggplot(mapping = aes(c(geo_sim))) + 
  geom_histogram(binwidth = 0.05) + 
  labs(x = "Spatial Similarity (Euclidean)", y = "Number of Samples")
ggplot(mapping = aes(c(env_sim))) + 
  geom_histogram(binwidth = 0.05) + 
  labs(x = "Environment Similarity (Euclidean)", y = "Number of Samples")

```

Community and geography are correlated and sites cluster around three locations
Also, geography uses up more of the similarity matrix space (see histograms)
Therefore, I coded geography as three factors (one for each cluster of sites)
Also, wetland/upland covaries with those three sites, but moisture content is
included in the environmental similarity matrix and should capture that variation

```{r fit_var_comp}

model <- varComp(lowk ~ taxon, varcov = list(com = com_sim, geo = geo_sim, env = env_sim) )
summary(model)

model <- varComp(lowk ~ taxon + all_data$geocode, varcov = list(com = com_sim, env = env_sim) )
summary(model)
levels(all_data$geocode)
all_data$geocode <- factor(all_data$geocode, levels(all_data$geocode)[c(2,3,1)])
model1 <- varComp(lowk ~ taxon + all_data$geocode, varcov = list(com = com_sim, env = env_sim) )
summary(model1)

```