### 2019-10-18 manuscript/manuscript_rev1.Rmd
Incorporate reviewer comments

---
title: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
output: 
    word_document:
        reference_docx: reference.docx
bibliography: library.bib
csl: philosophical-transactions-of-the-royal-society-b.csl 
---

```{r setup, include=FALSE}

library(tidyverse)
library(knitr)
library(flextable)
opts_chunk$set(echo = FALSE)

```

# To Do

 - [ ] Soften the language surrounding what progress we have made in microbial
   BEF - currently it is too harsh. Reference some of the newer
   multi-functionality literature.
 - [ ] Add an explanation of what figure two represents.
 - [ ] Reframe the paper about how there is ample evidence that underground
   biodiversity might play a role in ecosystem functioning. Add the caveat that
   current approaches to incorporating microbes into earth system models don't
   provide much increased variance explained. Use this to pitch an alternative
   approach: learning from geno-pheno mapping to angostically search for taxa
   that may regulate function upstream of the actual process being studied.

# Abstract

Microorganisms mediate many important ecosystem functions, yet it remains
unclear to what extent microbial diversity or community composition is
important for determining the rates of ecosystem-scale functions. This
uncertainty limits our ability to predict and manage crucial microbially
mediated processes, such as nutrient loss and greenhouse gas emissions. Our
lack of understanding stems from the relatively large diversity of
microorganisms, the difficulty in directly identifying functional groups, and
our limited ability to manipulate microbial community attributes. For this
reason, we propose that integrating traditional biodiversity-ecosystem function
research with ideas from genotype-phenotype mapping could provide the new
perspective our discipline needs. We identify two insights from
genotype-phenotype mapping that could be useful for microbial
biodiversity-ecosystem function studies: the concept of “agnostic” mapping and
the incorporation of covariates into models of ecosystem function to identify
microorganisms that uniquely contribute to ecosystem function. We illustrate
the potential for these approaches to elucidate microbial
biodiversity-ecosystem function relationships by analyzing a subset of
published data measuring methane oxidation rates from incubations of tropical
soil. We assert that combining the approaches of traditional
biodiversity-ecosystem function research with ideas from genotype-phenotype
mapping will not only generate novel hypotheses about how complex microbial
communities drive ecosystem function, but also help scientists predict and
manage changes to ecosystem functions resulting from human activities.

## Introduction

Ecology is broadly focused on understanding what creates the biodiversity we
observe and how that biodiversity shapes the ecosystems that humans depend
on. Earth's microbiome represents an important microcosm for that general
pursuit as most of the major biogeochemical cycles are mediated by
microorganisms. We have shown that certain ecosystem functions are positively
correlated with microbial biodiversity. However, attempts to understand the
mechanism underlying that relationship have borne little fruit. We argue that
to overcome this challenge we should look for inspiration from population
genetics and the process of genotype-phenotype mapping. In particular, we should look
more agnostically at microbial
communities to identify taxa that regulate ecosystem-scale functions. In
addition, we should attempt to understand the complex covariance that exists
between microbial communities, biogeographic patterns, and the underlying abiotic
environment. Finally, we reframe the question "Is microbial biodiversity
related to ecosystem function?" as "What is the unique contribution of the
microbiome to ecosystem function independent of the environment?"

### There is evidence that microbial diversity matters

There is evidence that microbial communities should matter for regulating
ecosystem function. Some studies have manipulated the connection between
environmental factors and microbial community composition through common garden
or reciprocal transplant experiments. These studies frequently report different
rates of ecosystem functions for different microbial communities under the same
environmental conditions [@reed2007]. This has been observed for plant
decomposition [@glassman2018, @strickland2009], plant phenology
[@panke-buisse2015], soil nitrogen cycling [@balser2005], and soil greenhouse
gas emissions [@cavigelli2000], among other functions. Therefore, variation in microbial
community composition can be associated with variation in ecosystem function.

Studies using a comparative approach rather than a manipulative one have also
demonstrated a link between microbial biodiversity and ecosystem function.  In
general, there is a significant correlation between belowground biodiversity
and multiple ecosystem functions [@jing2015, @delgado-baquerizo2016,
@delgado-baquerizo2017, @wagg2014]. It is unclear whether the positive
diversity-function relationship is due to the independent effect of variation
in community composition on ecosystem function or due to environmental
conditions that select for both diverse communities and high rates of ecosystem
function. Determining why we see this diversity-function relationship depends
on identifying the components of the community that are related to ecosystem
function.

### We need a new perspective on microbial biodiversity-ecosystem function relationships

Most sudies of microbial community function focus on either functional gene
abundance or community diversity.
In the case of functional genes, qPCR or shotgun metagenomic sequencing is used to
estimate the abundance of a gene or transcript that is a putative marker for a
microbial process (and thus a marker for the functional group that performs
that process). For example, the gene _mcrA_, which encodes a subunit of the
enzyme that performs the final step in methanogenesis, is commonly used as a
marker for methanogenesis and for the methanogen functional group. Other
examples include _pmoA_ for methanotrophy, _nifH_ for nitrification, and _nosZ_ for
denitrification. The abundance of these markers is often hypothesized to
predict the rate of their associated processes (for example, the
abundance of _mcrA_ is expected to relate to the rate of
methanogenesis). There are examples where this relationship is present
[@freitag2009, @freitag2010, @schnyder2018].  However, a review of such studies
found that the abundance of a functional gene or transcript is rarely
correlated with the rate of the corresponding process and depends on the
function of interest, with most effects either negative or not significant
[@rocca2015].

The second aspect of microbial community structure hypothesized to predict
ecosystem function is taxonomic or functional diversity. Diversity is either
estimated from sequence variants of a barcode gene such as the 16S rRNA gene or
manipulated through some proxy of diversity such as sequential dilution or
varying filter sizes. Studies that experimentally manipulate diversity
generally find a relationship between the applied diversity treatment and
ecosystem function, including methanotrophy [@schnyder2018], phosphorus
leaching [@wagg2014], greenhouse gas emissions [@wagg2014], decomposition
[@maron2018], and nitrogen cycling [@philippot2013]. However, in many cases
richness is confounded with other factors such as the presence or absence of
major phylogenetic groups (nematodes, fungi) or abundance of microbial cells.
In addition, assembled microbial communities risk confounding diversity with
composition, for example by more likely including the most productive taxa in
the highest diversity treatment. In general, the relationship between microbial
diversity is positive and significant, but adds little variance over abiotic
conditions or macroorganismal diversity [@graham2016, @jing2015].

Despite the positive diversity-function relationship for microbes, models that
incorporate microbial features only marginally improve predictions over
environmental variables [@graham2016]. In addition, we've made little progress
in identifyng the components of microbial communities that drive these
diversity-function relationships [@rocca2015].
Overall, functional gene abundance and community diversity improve models of
ecosystem function less than one third of the time and increase variance
explained by an average of only 8 percentage points [@graham2016].

Up to this point, microbial biodiversity-ecosystem function research has
followed a similar path to macrobial biodiversity-ecosystem function research,
particularly compared to plant communities. However, there are important
differences between plants and microbes that might inform how we pursue this
research going forward. For example, microbial communities are more diverse and
contain greater numbers of individuals than plant communities and are thus more
difficult to sample comprehensively. Unlike plants, functional groups of
microbes are rarely determined through observation or direct measurement, but
rather must be indirectly inferred based on taxonomic assignment of marker
genes (such as 16S rRNA genes), by sequencing functional genes from
metagenomes, or by quantifying specific functional genes using qPCR.  This adds
uncertainty in our ability to estimate the size of functional groups.  The
ability to experimentally manipulate diversity or community composition is much
more limited for microbes. Given all of these differences (and many others),
the challenge of linking microbial communities to ecosystem functions may be
less like that of plant communities and more analogous to the complex task of
linking genomic variation to organismal phenotypes. For this reason, we propose
that integrating traditional biodiversity-ecosystem function research with
ideas from genotype-phenotype mapping could provide the new perspective our
discipline needs.

### Genotype-phenotype as a source of inspiration

Given the relative lack of success to date, it is time to rethink how we
approach the challenge of mapping microbial community structure to ecosystem
function. Microbial ecologists are not the only biologists who are attempting
to determine the relationship between a complex set of highly-variable data and
an aggregate function. This kind of “many-to-one” mapping is analogous to the
challenge of identifying the genetic basis of complex traits in organismal
populations. In such genotype-phenotype mapping studies, a population
exhibits variation in a phenotype (e.g., height or disease state) as well as
variation in thousands of genetic loci or single nucleotide polymorphisms (SNPs).
To identify the genetic basis for a trait, investigators sample from this
population and correlate phenotype with genotype. In most cases, there are many
more loci than individuals and we do not know whether the SNPs are causally
linked or are simply in linkage disequilibrium with a causal mutation.

There are a number of parallels between the challenge faced by organismal
biologists and that facing microbial community ecologists. They are both
“many-to-one” mapping challenges which involve large numbers of statistical
comparisons. Both are attempting to identify causal relationships that are
potentially confounded by complex patterns of covariation. For example,
with local selection or distance-decay. Further, a positive association may
either indicate a causal relationship or simply an association between a
significant taxon and a causal taxon (i.e., microbial linkage
disequilibrium). There is often no strong expectation about which entities
(i.e., which genomic regions or which microbial taxa) are most likely to be
causally-related to phenotype or function, and thus “agnostic” approaches are
needed.  Both are sensitive to exactly how the “mapping” question is asked, and
both ultimately require manipulation (of genes or taxa) to establish causation. We
describe each of these parallels below in more detail and provide an example
of how these ideas could be applied to microbial data.

### The importance of a taxonomically “agnostic” approach

[This is super redundant with the previous section.]

The traditional approach to macroorganismal biodiversity-ecosystem function
research is to measure or manipulate the diversity of a taxonomic group (e.g.,
plants) and look for an association with the function performed by that group
(e.g., primary productivity). In the broadest sense, we can think of plants as
a “functional group,” i.e., a group of taxa united by their ability to perform
a particular ecosystem function. Ecologists may further divide a functional
group (e.g., plants) into smaller functional groupings (e.g., forbs), defined
by the details of how they perform their particular ecosystem function.

However, for microbes our knowledge of functional groups is much more limited.
From a very limited number of cultured isolates we have a provisional
understanding of which microbes might be involved in some ecosystem functions.
And by sequencing the genomes of these isolates, we have identified genetic
markers for certain functions. But most microbial taxa remain uncultured and we
do not know the function of most microbial taxa detected in environmental
samples [@hug2016, @martiny2019, @steen2019]. In addition, there have been
recent discoveries of functions in unexpected taxonomic groups, for example
methanogenesis by fungi and cyanobacteria, a function previously considered
restricted to the Archaea [@bizic-ionescu2019, @lenhart2012]. Therefore,
estimating microbial functional group abundance can be challenging.

Beyond, for example, identifying previously undocumented methanotrophs that are
related to methane oxidation, it's possible that more complex ecological
regulatory patterns exist in microbial communities. In other words, it may
be that methanotroph abundance does not limit methanotrophy in a given ecosystem.
[Something about metabolic networks...]
Instead, the rate
of methane flux to the atmosphere could be controlled by methanotrophs,
methanogens, bacteria that produce the fermentative byproducts that methanogens
depend on, nutrient cyclers that regulate the availability of nitrogen or
phosporus, or even grazers and other predators or pathogens. In other words, the regulation of
ecosystem function by microbial communities could be direct or indirect, much
like genetic regulation within an organism.

Because of this, it would be prudent to look more agnostically at microbial
communities to identify taxa or groups of taxa that are important for
predicting the rate of ecosystem functions rather than assuming that the
genetic markers we have provisionally identified for a given function represent
the most likely taxa involved. This agnostic approach is analogous to the
approach of many genotype-phenotype mapping studies (e.g., genome-wide
association studies or GWAS), which often look for associations between a
phenotype and loci anywhere in a genome.

### Controlling for population stratification

It is widely accepted that organisms, including microorganisms, exhibit
population stratification due to geographic and environmental separation
[@martiny2006, @wright1943]. Genome-wide association studies generally control
for population stratification using shared ancestry among cases and controls when
modeling the connection between genotype and phenotype. The classic example is
the latitudinal gradient of both height and genotypic similarity in Europe,
which results in spurious associations between human height and genetic
variation [@berg2019, @novembre2008, @sohail2019]. Perhaps a more illustrative
example is the correlation between chopstick usage and certain genetic markers
in Asian populations, which is unlikely to be due to an actual genetic basis
for chopstick skills. Spurious associations arise
because of heterogeneity in population structure where any genetic variant in a
population with an extreme phenotype will tend to be associated with that
phenotype whether it is related or not.

Genome-wide association studies correct for stratification using a variety of methods.
Generally, they ignore the underlying environmental and spatial
distance between samples and instead use shared ancestry as a proxy for local
selection and assortative mating.  A common approach is to perform a regression
of phenotype and shared ancestry (computed as the first one or more principal
components of the genotype matrix) and then use the residuals from this model
as the values for phenotype in a
subsequent regression using the genotypes directly [@price2006]. This principal
component analysis is designed to test the effect of individual genes after
removing shared ancestry among individuals. Another approach, employed in our
example, is variance component modeling (or mixed modeling, heirarchical
modeling, etc.) where genotypic similarity is included as a covariate in the
model [@kang2010].

While these represent two common statistical approaches to
control for population stratification, it is also considered good practice to
control for stratification experimentally. This can be accomplished by 
selectng a study population that is relatively homogeneous. Otherwise, even
large, well-replicated GWAS studies can result in
false positives. For example, it was recently revealed that GWAS studies on
height of Europeans failed to adequately control for stratification. When their
results were confirmed using a similarly sized, but more homogeneous study
population, many of the genetic markers for height were no longer significant [@berg2019,
@sohail2019].

[This paragraph needs work]

Typically, microbial biodiversity-ecosystem function studies do not account for
population stratification (i.e., community similarity across samples), although
there are some exceptions [@meyer2019, @qin2012, @lloyd-price2019].  However,
community similarity (the community analogue of shared ancestry among
organisms) is not as tightly linked to geography or environment as is shared
ancestry among organisms, and it could be very useful to account for these
separately in microbial studies, especially if one is particularly interested
in how composition alters function independent of the underlying environmental
variation.

### What might this look like/More about the analogy

The relationship between structure and function could be complex and indirect.
i.e. not just methanogens making methane, but metabolic and regulatory networks

A significant correlation between a taxon and a function could mean a number of
things.
1. That microbe is performing that function right now and its abundance is
   related to function
2.

This whole compositional data/relative abundance thing also seems problematic
for correlating abundance with function. Really what we're saying is "this read is differentially abundant in low vs. high flux ecosystems." Not abundance is _correlated_ with flux.
1. It's not **abundance**.
2. An increase in relative abundance in one taxon implies nothing about the
   abundance of that taxon

   wootincrease in relative abundance in one taxon implies nothing about the
abun
3. It is perhaps somewhat more likely that

### We need a new perspective on microbial community-ecosystem function relationships

Many of the topics in this first section are covered later (e.g. in diversity
matters). What is the purpose of this section - how does it move the story
along? It should motivate this work and put it in a larger context. This is the
hook. Maybe the best way to do this is introduce what we know, then create a
conflict by showing what we don't know, and then give a brief read map of how
we're going to fix that. Possibly send these 2-3 paragraphs to Bebo. Maybe also
read some writing advice on hooks (maybe in Schimel's book).

Ecologists have long investigated the effects of changing biodiversity on
ecosystem function, documenting, for example, relationships between terrestrial
plant community richness and primary productivity [@hooper2005]. The study of
the relationship between microbial biodiversity and ecosystem functions is much
more recent and has generated mixed results [@schimel1998, @singh2010]. Despite
the fact that microbes mediate many important ecosystem functions, it remains
maddeningly unclear to what extent microbial diversity or community composition
is important for determining the rates of ecosystem-scale functions.  For
example, numerous studies have attempted to correlate microbial functional
group abundance or diversity with the rate of various ecosystem functions and
most are unsuccessful [@rocca2015, @graham2016]. This uncertainty limits our
ability to predict and manage crucial microbially mediated ecosystem functions,
such as nutrient loss and greenhouse gas emissions. As a discipline, we need a
new perspective on these questions.

Numerous studies have attempted to link microbial diversity or functional group
abundance  to ecosystem function. In general, there is a significant correlation between
belowground biodiversity and multiple ecosystem functions [@jing2015,
@delgado-baquerizo2016, @delgado-baquerizo2017, @wagg2014]. However, models that incorporate microbial features
only marginally improve predictions overall environmental variables [@graham2016]. In addition, we've made
little progress in identifyng the components of microbial communities that
drive these diversity-function relationships [@rocca2015]. It is unclear
whether the positive diversity-function relationship is due to the independent
effect of variation in community composition on ecosystem function or due to
environmental conditions that select for both diverse communities and high
rates of ecosystem function. Determining why we see this diversity-function
relationship depends on identifying the components of the community that are
related to ecosystem function independent of the environmental variation.

I don't love the below paragraph

Our lack of understanding stems from many sources. Microbial communities are
more diverse and contain greater numbers of individuals than plant communities
and are thus more
difficult to sample comprehensively. Unlike plants, functional groups of
microbes are rarely determined through observation or direct measurement, but
rather must be indirectly inferred based on taxonomic assignment of marker
genes (such as 16S rRNA genes), by sequencing functional genes
from metagenomes, or by quantifying specific functional genes using qPCR. This adds uncertainty in our ability to estimate the size of functional groups.
The ability to experimentally manipulate microbial diversity or community
composition is much more limited for microbes. Given all of these differences
(and many others), the challenge of linking microbial communities to ecosystem
functions may be less like that of plant communities, and more analogous to the
complex task of linking genomic variation to organismal phenotypes. For this
reason, we propose that integrating traditional biodiversity-ecosystem function
research with ideas from genotype-phenotype mapping could provide the new
perspective our discipline needs.

### There is evidence that microbial diversity matters

Perhaps remove this paragraph. Maybe add an explanation of "independent of the
environment" slash "uniquely attributed"

Perhaps our uncertainty arises because there is no general relationship between
microbial biodiversity and any particular ecosystem function. For example, if
local selection always optimizes the available microbial biodiversity to
maximize ecosystem function, then microbial community composition should not
matter for predicting the rates of microbially mediated functions. Instead, the
rates of these functions should be determined by the underlying environmental
variation. In this case, the microbial community would simply act as a conduit
through which the abiotic environment alters ecosystem function. Attempts to
document a general relationship between microbial community attributes and
function would then fail (since the primary drivers are environmental).

However, there is evidence that this is not always the case [super awkward]. Some studies have
manipulated the connection between environmental factors and
microbial community composition through “common garden” or reciprocal
transplant experiments, and they frequently report different rates of ecosystem
functions for different microbial communities under the same environmental
conditions [@reed2007].  This has been observed for plant decomposition
[@glassman2018, @strickland2009], plant phenology [@panke-buisse2015], soil
nitrogen cycling [@balser2005], and soil greenhouse gas emissions
[@cavigelli2000], among others. Therefore, variation in microbial community
composition can be associated with variation in ecosystem function, independent
of environmental variation. So why has this been difficult to consistently
document in the field?

### Traditional approaches to quantifying this relationship have provided only minor improvements

Most comparative studies of microbial community function in the field focus on
one of two aspects of microbial community structure that are hypothesized to
predict ecosystem function. The first aspect is “functional” gene or transcript
abundance.  In this case, qPCR or shotgun metagenomic sequencing is used to
estimate the abundance of a gene or transcript that is a putative marker for a
microbial process (and thus a marker for the functional group that performs
this process). For example, the gene mcrA, which encodes a subunit of the
enzyme that performs the final step in methanogenesis, is commonly used as a
marker for methanogenesis and for the methanogen functional group. Other
examples include pmoA and methanotrophy, nifH and nitrification, and nosZ and
denitrification. The abundance of these markers is often hypothesized to be
predictive of the rate of their associated processes (for example, the
abundance of mcrA is often hypothesized to be related to the rate of
methanogenesis). There are examples where this relationship is present
[@freitag2009, @freitag2010, @schnyder2018].  However, a review of such studies
found that the abundance of a functional gene or transcript is rarely
correlated with the rate of the corresponding process and depends on the
function of interest, with most effects either negative or not significant
[@rocca2015].

The second aspect of microbial community structure hypothesized to predict
ecosystem function is taxonomic or functional diversity. Diversity is either
estimated from sequence variants of a barcode gene such as the 16S rRNA gene or
manipulated through some proxy of diversity such as sequential dilution or
varying filter sizes. Studies that experimentally manipulate diversity
generally find a relationship between the applied diversity treatment and
ecosystem functions, including for methanotrophy [@schnyder2018], phosphorus
leaching [@wagg2014], greenhouse gas emissions [@wagg2014], decomposition
[@maron2018], and nitrogen cycling [@philippot2013]. However, in many cases,
richness is confounded with other factors such as the presence or absence of
major phylogenetic groups (nematodes, fungi) or abundance of microbial cells.
In addition, assembled microbial communities pose the same challenges as
macroorganismal diversity experiments, for example the highest diversity
treatment will more likely contain the most productive taxon. A review of
microbial diversity studies shows that microbial community taxonomic and
functional diversity add little explanatory power to models of ecosystem
function [@graham2016]. Overall, functional gene abundance and community
diversity improve models of ecosystem function less than one third of the time
and increase variance explained by an average of only 8 percentage points
[@graham2016].

### Genotype-phenotype mapping as a source of inspiration

Given the relative lack of success to date, it is time to rethink how we
approach the challenge of mapping microbial community structure to ecosystem
function.  Microbial ecologists are not the only biologists who are attempting
to determine the relationship between a complex set of highly-variable data and
an aggregate function.  This kind of “many-to-one” mapping is analogous to the
challenge of identifying the genetic basis of complex traits in organismal
populations. In such “genotype-phenotype” mapping studies, a population
exhibits variation in a phenotype (e.g., height or disease state) as well as
variation in potentially thousands of single nucleotide polymorphisms (SNPs).
To identify the genetic basis for a trait, investigators sample from this
population and correlate phenotype with genotype. In most cases, there are many
more loci than individuals and we do not know whether the SNPs are causally
linked or are simply in linkage disequilibrium with a causal mutation.

There are a number of parallels between this challenge faced by organismal
biologists and that facing microbial community ecologists.  They are both
“many-to-one” mapping challenges, which involve large numbers of statistical
comparisons.  Both are attempting to identify causal relationships that are
potentially confounded by very complex patterns of covariation.  There is often
no strong expectation about which entities (i.e., which genomic regions or which
microbial taxa) are most likely to be causally-related to phenotype or
function, and thus “agnostic” approaches are needed.  Both are sensitive to
exactly how the “mapping” question is asked, and ultimately require
manipulation (of genes or taxa) to establish causation.  We describe each of
these parallels below in more detail, and provide an example of how these ideas
could be applied to microbial data.

### The importance of a taxonomically “agnostic” approach

The traditional approach to microorganismal biodiversity-ecosystem function
research is to measure or manipulate the diversity of a taxonomic group (e.g.,
plants) and look for an association with the function performed by that group
(e.g., primary productivity).  In the broadest sense, we can think of plants as
a “functional group,” i.e., a group of taxa united by their ability to perform
a particular ecosystem function. Ecologists may further divide a functional
group (e.g., plants) into smaller functional groupings (e.g., forbs), defined
by the details of how they perform their particular ecosystem function.

However, for microbes our knowledge of functional groups is much more limited.
From a very limited number of cultured isolates we have a provisional
understanding of which microbes might be involved in some ecosystem functions.
And by sequencing the genomes of these isolates, we have identified genetic
markers for certain functions. But most microbial taxa remain uncultured and we
do not know the function of most microbial taxa detected in environmental
samples [@hug2016] (@martiny2019 or @steen2019). In addition, there have been
recent discoveries
of functions in unexpected taxonomic groups, for example methanogenesis by
fungi and cyanobacteria, a function previously considered restricted to the
Archaea [@bizic-ionescu2019, @lenhart2012].

Beyond, for example, identifying previously undocumented methanotrophs that are
related to methane oxidation, it's possible that more complex ecological
regulatory patterns exist in microbial communities. In other words, it may not
be methanotrophs limiting methanotrophy in a given ecosystem. Instead, the rate
of methane flux to the atmosphere could be controlled by methanotrophs,
methanogens, bacteria that produce the fermentative byproducts that methanogens
depend on, nutrient cyclers that regulate the availability of nitrogen or
phosporus, or even grazers and other predators or pathogens. The regulation of ecosystem function by microbial communities could be direct or indirect.

Because of this, it would be
prudent to look more agnostically at microbial communities to identify taxa or
groups of taxa that are important for predicting the rate of ecosystem
functions rather than assuming that the genetic markers we have provisionally
identified for a given function represent the most likely taxa involved.  This
agnostic approach is analogous to the approach of many genotype-phenotype
mapping studies (e.g., genome-wide association studies or GWAS), which often
look for associations between a phenotype and loci anywhere in a genome.

### Dealing with multiple comparisons

Microbial biodiversity-ecosystem function studies and genotype-phenotype
mapping studies are both “many-to-one” mapping challenges, which involve large
numbers of statistical tests.  Hundreds to thousands of tests are routinely
made per study, greatly inflating the number of false positives identified
using statistical hypothesis testing approaches.  Some microbial
biodiversity-ecosystem function studies do not correct for this, while others
use approaches that may unnecessarily inflate the false negative rate, such as
the Bonferroni correction. The Bonferroni correction has been widely considered
to be too conservative, particular for exploratory studies designed to generate
hypotheses [@perneger1998, @rothman1990, @thomas1985]. Statisticians have
developed a number of less-conservative approaches to correct for multiple
comparisons by controlling the false discovery rate in order to balance the
tradeoff between Type I and Type II errors [@benjamini1995, @storey2002].
These approaches are commonly used in genotype-phenotype mapping studies.

### Controlling for population stratification

Maybe introduce "uniquely associated" here.

It is widely accepted that organisms, including microorganisms, exhibit
population stratification due to geographic and environmental separation
[@martiny2006, @wright1943]. Genome-wide association studies generally account
for population structure due to shared ancestry among cases and controls when
modeling the connection between genotype and phenotype. The classic example is
the latitudinal gradient of both height and genotypic similarity in Europe,
which results in spurious associations between human height and genetic
variation [@berg2019, @novembre2008, @sohail2019]. Spurious associations arise
because of heterogeneity in population structure where any genetic variant in a
population with an extreme phenotype will tend to be associated with that
phenotype whether it is related or not. To correct for this
covariance structure, GWAS models incorporate genotypic similarity to correct
for shared ancestry using a variety of methods. A common example is to correct
for population structure by performing regression of the first one or more
principal components of genotype on both genotype and phenotype. Then, the
residuals from this model are used as the new genotype and phenotype with
structure removed [@price2006]. Another approach, employed in our example, is
variance component
modeling (or mixed modeling, heirarchical modeling, etc.) where genotypic
similarity is included as a covariate in the model as a random
effect [@kang2010]. While these represent two common statistical approaches to
control for population stratification, it is also considered good practice to
select a study population that is relatively homogeneous when designing an
experiment. Otherwise, even large, well-replicated GWAS studies can result in
false positives, as recently revealed by two studies reversing conclusions from
previous studies of the relationship between genetics and human height
[@berg2019, @sohail2019].

Typically, microbial biodiversity-ecosystem
function studies do not account for population stratification (i.e., community
similarity across samples), although there are some exceptions [@meyer2019,
@qin2012]. Genome-wide association studies generally ignore the underlying
environmental and spatial distance between samples and instead uses shared
ancestry as a proxy for these variables.  However, community similarity (the
community analogue of shared ancestry among organisms) is not as tightly linked
to geography or environment as is shared ancestry among organisms, and it could
be very useful to account for these separately in microbial studies, especially
if one is particularly interested in how composition alters function
independent of the underlying environmental variation.

## An example: high-affinity methane oxidation

To illustrate the ideas outlined above, we reanalyzed a subset previously
published data [@meyer2019] using a modified procedure from the original
version.  A full description of the study design, samples, and data generation
can be found in that manuscript. Briefly, these data were gathered from intact
soil cores taken from diverse ecosystems of the Congo Basin in Gabon, Africa.
Cores were incubated in the laboratory under different concentrations of
methane to identify the rates of specific methane cycling pathways. For this
example, we will analyze data from just one of these pathways, high-affinity
methane oxidation (the oxidation of atmospheric concentrations of methane),
which we will refer to below as “methane oxidation”. In addition, for
simplicity we only include amplicon sequences from the DNA-inferred community
and not the RNA-inferred community, both of which are presented in the original
paper [@meyer2019].  The data we analyzed include methane oxidation
measurements, amplicon sequence variants (ASVs; @callahan2017) generated using
DADA2 and inferred from 16S rRNA gene sequences, pmoA abundance estimates (via
qPCR), latitude and longitude, and four environmental covariates (soil
moisture, bulk density, carbon, and nitrogen).

Analyses were conducted in the R statistical environment using the phyloseq
package [@mcmurdie2013, @rcoreteam2019].  The relative abundances of ASVs were
corrected using the variance stabilizing transformation from DESeq2 [@love2014,
@mcmurdie2014]. We first test typical measures of microbial structure including
functional gene abundance and community richness, which was estimated using the
breakaway package [@willis2015]. We then demonstrate significant covariation
between community structure (estimated as Bray-Curtis distance using vegan),
environmental variation (euclidean distance), and geographic distance
(euclidean distance) using Mantel tests [@bray1957, @oksanen2019].  Finally, we
present one approach to identifying taxa which are significantly associated
with function independent of the environment by fitting variance component
models using varComp to test the relationship between relative abundance of
each ASV and methane oxidation rate [@qu2013]. To illustrate how including
different covariates (environmental, geographic, and community) can result in
different conclusions about which taxa are associated with function, we fit
this model with and without random effects variance components for
environmental similarity, geographic site ID, and Bray-Curtis similarity.
Significant taxa were determined by controlling the false discovery rate at
q-value < 0.05 [@storey2002]. Figures were created using ggplot2
[@wickham2019].

## Results & Discussion

Methane oxidation rate was not significantly correlated with the abundance of
methanotrophs (inferred from _pmoA_ gene
abundance) or 16S read richness (Table 1, Figure 1). We tested collinearity between each
pair of distance matrices for community, environment, and geography using
Mantel tests and estimated p-values by permutation. We found a moderate and
significant correlation between community composition and environmental
variation, geography and community composition, and geography and environmental
variation (Table 2, Figure 2). Principal coordinate plots show that beta
diversity of samples separated by site ID and by ecosystem type (wetland or
upland; Figure 2). This indicates substantial spatial and environmental
structure for microbial populations which should be controlled for to
rigorously identify associations between taxa and function. Finally, we tested
the effect of the relative abundance of
each ASV on methane oxidation rate. We found different numbers of taxa
significantly associated with methane oxidation depending on which covariates
were included in the model (Table 3). After correcting for the false discovery
rate, 460 unique 16S reads were identified as significantly correlated with function.
In the model that controlled for community structure using all three covariates
(community, environment, and geography), only 6 of these reads were found to be
significiantly correlated with methane oxidation rate.
Though we cannot infer function from 16S sequences, these 6 taxa fall into
three genera and one class with cultured representatives that are not known to
consume methane [@belova2009, @chang2011, @domeignoz-horta2019, @fritz2004].
Their effect on methane oxidation rate ranges from 0.5 to 1.5 (Figure 3). (The
slope of the relationship between relative abundance and methane oxidation
rate.) An
effect of 1.5 means that for a one unit increase in relative abundance there is
a 1.5 increase in k (the rate of exponential decrease in methane concentration
over time).

In this paper, we argue that the traditional approach to microbial
biodiversity-ecosystem function research often ignores the complexity of
microbial community dynamics, in particular the complex patterns of covariation
that often arise among microbial community similarity, environmental
similarity, and spatial proximity.  We propose that a better approach would be
to learn from other complex many-to-one mapping problems in biology,
particularly genotype-phenotype mapping.  These studies often explicitly
account for covariation, they correct for multiple comparisons in powerful
ways, and they frequently take an “agnostic approach” that does not assume a
particular relationship between structure and function.

We illustrate the potential for these approaches to elucidate microbial
biodiversity-ecosystem function relationships by analyzing a subset of
published data from incubations of tropical soil.  In this example, soil cores
from different ecosystems exhibited different rates of methane oxidation, but
simple correlations between these rates and the diversity or abundance of
putative methane oxidizing bacteria were not informative.  These results agree
with recent reviews of the literature that demonstrate that such simple
approaches are often not a fruitful avenue for elucidating microbial
structure-function connections.

[redundant with the former results section]

We then asked if the relative abundance of particular taxa were related to
ecosystem function, and we identified 460 taxa whose relative abundances were
associated with rates of methane oxidation after controlling the false
discovery rate (q-value < 0.05).  These taxa could be related to ecosystem
function in multiple ways.  The most interesting possibility is that each of
these taxa is statistically related because it is causally connected to the
function. This could be direct, for example an organism that consumes methane,
or indirect, for example an organism that regulates substrates necessary for
consumers of methane. In either of these cases, the taxon could be useful as a
biomarker of function or as an organism to investigate in order to better
understand the biological drivers of variation in methane oxidation.

Alternatively, a significant association could occur for non-causal reasons.
For example, any organism that tends to be in high abundance where methane
oxidation rates are high would be correlated with methane oxidation, even if it
has no causal relationship. This could be because such an organism is favored
under the same environmental conditions that favor methane oxidizing bacteria
(or that favor methane oxidation in general).  Such covariation can drive
associations that are not causal, but the effects of such covariation can be
reduced through a number of approaches (many developed by biologists interested
in genotype-phenotype mapping).

In our example, we showed that the abundances of microbial taxa exhibit complex
patterns of covariation with each other and with environmental conditions and
spatial location. These biogeographic patterns might be even stronger if we
could sample the populations more intensively [@meyer2018]. Once we account for this
covariation, our list of taxa associated with function was reduced to 6.  These
6 taxa include taxa that are not known to directly contribute to methane
oxidation, suggesting that the drivers of methane oxidation may be indirect,
perhaps mediated by ecological interactions among taxa from multiple functional
groups.

This approach is powerful because it accounts for the covariances that arise
from the fundamental ecological processes that drive community assembly and
underlie biogeographical patterns [@martiny2006, @vellend2010]. It is also powerful because it is
focused on a much more specific version of the “biodiversity-ecosystem
function” question, i.e., it asks “which taxa are uniquely associated with
function?” By “uniquely associated,” we mean those taxa associated with
function irrespective of environmental conditions, local community structure,
or spatial proximity.  This is not only a more specific question than is
usually asked in microbial biodiversity-ecosystem function studies, but it is
also a more appropriate one, especially if one is interested in how to
incorporate microbial community data into ecosystem models. These models depend
on identifying taxa that add explanatory power
beyond that provided by other factors, such as environmental conditions.

We identified three insights from genotype-phenotype mapping that could be
useful for microbial biodiversity-ecosystem function studies: the concept of
“agnostic” mapping, the use of more powerful ways to account for multiple
comparisons, and the incorporation of covariates.  But there are other insights
to be gained as well.  For example, some phenotypes (e.g., the propensity for
diseases such as Parkinson's) are controlled by a single genetic locus [@kerem1989, @macdonald1992].
However, most phenotypes studied to date arise from the influence of many loci
of small effect (as well as environmental factors; [@hindorff2009, reich2001]). Organismal
biologists have developed approaches tailored to identifying such loci; for
example, by sampling organismal populations in a way that constrains genetic
variation unrelated to phenotype. Similarly, for some ecosystem functions, it
is possible that a single microbial taxon could substantially influence its
rate. For example, methane flux from permafrost in Sweden may be controlled by
a single taxon [@mccalley2014]. However, most microbially mediated ecosystem functions are
likely the result of interactions (direct and indirect) among many taxa.  We
could maximize our ability to identify taxa of small effect by constraining
variation, e.g., by sampling microbial communities in a way that makes them
similar in structure (similar ecosystem, soil type, and abiotic conditions)
while still varying in function.

Ultimately, the relationships identified in genotype-phenotype mapping studies
must be verified.  There are multiple ways that this verification is
accomplished. In some cases, organisms can be artificially selected for a
particular phenotype (e.g., through experimental evolution in an environment
that favors the phenotype of interest) and the genetic changes that occur in
response to selection can be compared to those identified via mapping studies
(such as GWAS).  An analogous approach for microbial biodiversity-ecosystem
function studies would be to apply artificial ecosystem selection (sensu @swenson2000)
on a given function and compare the taxa that change in response to selection
with those identified via a comparative approach (such as the one illustrated
in our example).

The most common way that loci identified in a mapping study are verified is
through manipulative genetics.  The identified loci can be knocked out or
over-expressed, and the effect on phenotype compared to that predicted from
mapping studies.  There is no direct analogue for this in microbial
biodiversity-ecosystem function studies.  In some cases it may be possible to
inhibit a particular functional group through the use of a specific
antimicrobial or a chemical inhibitor [@maxson2016], but this is not generally true.  It
may be possible in some cases to isolate a microorganism of interest in pure
culture and add it back to a particular ecosystem, transiently increasing its
abundance (roughly analogous to “overexpressing” a gene).  Synthetic
communities (contrived assemblages of microorganisms) may ultimately be the
most powerful way to test hypotheses about microbial biodiversity-ecosystem
function relationships, but currently these approaches are limited by the small
number of taxa that can be routinely cultured from most environments.

## Conclusion

Microbial biodiversity-ecosystem function research has demonstrated that
functional group abundance (measured by qPCR of functional genes) is often not
a good predictor of ecosystem function and that microbial
diversity metrics on average do not add much power to ecosystem models.  A new
perspective on how to determine the relationship between microbial communities
and ecosystem functions is sorely needed.  Organismal biologists have over a
hundred years of experience identifying relationships between complex sets of
highly-variable data (genotypes or genome sequences) and aggregate functions
(organismal phenotypes).  We assert that combining the approaches of
traditional biodiversity-ecosystem function research with ideas from
genotype-phenotype mapping could provide this new perspective.  This
integration could not only make underutilized approaches such as covariate
modeling and artificial selection more available to microbial ecologists, but
also provide instructive examples of how best to conceive of microbial
biodiversity-ecosystem function questions.  If this integration is successful,
it is possible that in the not-so-distant future our field will be able to
robustly identify taxa, genes, or even molecules that will allow us to
accurately predict the response of ecosystems to environmental change. Doing so
will not only generate novel hypotheses about how complex microbial communities
drive ecosystem function, but also help scientists predict and manage changes
to ecosystem functions resulting from human activities.

# Acknowledgments

This project was supported by the National Science Foundation – Dimensions of
Biodiversity program (DEB 14422214), the National Science Foundation Graduate
Research Fellowship Program (DGE 1255832), and the ARCS Foundation Florence and
Mike Nudelman Scholarship. The data re-analyzed here were generated with
permission of the Government of Gabon, Centre National de la Recherche
Scientifique et Technologique (Permit No AR0035/14/MESRS/CENAREST/CG/CST/CSAR)
and the assistance of F. Bivigou, H. Memiaghe, L. Tchignoumba, E.  Tobi, and I.
Akendengue. The University of Oregon, the Gabon-Oregon Transnational Center on
Environment and Development, the Smithsonian Conservation Biology Institute,
and Shell Gabon provided financial and logistical support for the original
study re-analyzed here. We also thank W. Cresko for insightful discussions on
population genetics and H.  Tavalire for contributions to the statistical
analyses.

# References

::: {#refs} 
:::

# Tables and Figures

```{r copy_num}
model_out <- read.csv('../output/model_out.csv')
flextable(model_out) %>% 
  compose(i = 1, j = 1, 
    value = as_paragraph(as_i("pmoA"), " copy number")) %>% 
  compose(part = "header", j = 4, 
    value = as_paragraph("t-statistic")) %>% 
  compose(part = "header", j = 5, 
    value = as_paragraph("p-value")) %>% 
  autofit()
```

Table 1. Functional gene abundance and ASV richness are not good predictors of
methane oxidation rate. Linear models predicting methane oxidation rate from
measures of microbial community structure.

```{r mantel}
mantel_table <- read.table('../output/mantel_table.tsv')
flextable(mantel_table) %>% 
  compose(part = "header", j = 2, 
    value = as_paragraph("Mantel statistics (r)")) %>% 
  compose(part = "header", j = 3, 
    value = as_paragraph("95% upper quantile of permutations")) %>% 
  compose(part = "header", j = 4, 
    value = as_paragraph("p-value")) %>% 
  autofit()
```

Table 2. Mantel tests for each pair of dissimilarity matrices. Community
distance matrix was based on Bray-Curtis distance while both environment and
geography distance matrices were based on Euclidean distance. P-values
determined by permutation test with 999 permutations.

```{r n_taxa}
n_taxa <- read.table('../output/comp_adds_removes.tsv')
flextable(n_taxa) %>% 
  compose(part = "header", j = 1, 
    value = as_paragraph("Terms")) %>% 
  compose(part = "header", j = 2, 
    value = as_paragraph("Removed")) %>% 
  compose(part = "header", j = 3, 
    value = as_paragraph("Added")) %>% 
  compose(part = "header", j = 4, 
    value = as_paragraph("Significant")) %>% 
  autofit()
```

Table 3. Number of significant taxa after including each set of covariates in a
variance component model. Removed and Added columns are relative to the
no-covariate model. Significance is determined by controlling the false
discovery rate at q-values < 0.05.

```{r methane, out.width="768px", out.height="364px"}
knitr::include_graphics('../figures/trad_cors.pdf')
```

Figure 1. Methane oxidation rate is not correlated with functional gene
abundance or ASV richness. Correlations between community attributes and
ecosystem function. A) Abundance of the functional gene pmoA (n = 42) and B)
ASV richness (n = 44). Lines represent the ordinary least squares regression
line with standard errors.

```{r structure, out.width="960px", out.height="1152px"}
knitr::include_graphics('../figures/pcoa_multi.pdf')
```

Figure 2. Microbial community composition is spatially and environmentally
structured. Principle coordinate plots of Bray-Curtis distance representing the
first three axes of community composition. In A, B, and C, points are
identified by Site ID and in D, E, and F, points are identified by wetland or
upland ecosystem. All four environmental covariates separate strongly by
wetland/upland. Axis length is proportional to variance explained as indicated
in parentheses. PC = principal coordinate.

```{r taxa, out.width="768px", out.height="384px"}
knitr::include_graphics('../figures/effect_sizes.pdf')
```

Figure 3. Taxa associated with methane oxidation rates after controlling for
geographic location, environmental similarity, and community composition.
Points are estimates for the linear relationship between the relative abundance
of a single ASV and methane oxidation rate with standard errors from variance
component models including similarity matrices as covariates for community and
environment and site ID for geographic location. Amplicon sequence variants are
labelled at the finest resolution available: genus for all except the Group 1
Acidobacterium. Points are identified by Phylum. Significant taxa were
determined by controlling the false discovery rate at q-value < 0.05.