### 2019-10-18 manuscript/manuscript_rev1b.Rmd

## Outline

### We need a new perspective on microbial community-ecosystem function relationships

### There is evidence the microbial diversity matters

### Current approaches to quantify this relationship have provided only minor improvements

### Genotype-phenotype as a source of inspiration

### A taxonomically "agnostic" approach

### Controlling for population stratification/the effect of the microbiome independent of the environment

## Intro

Ecology is broadly focused on understanding what creates the biodiversity we
observe and how that biodiversity shapes the ecosystems that humans depend
on. Earth's microbiome represents an important microcosm for that general
pursuit as most of the major biogeochemical cycles are mediated by
microorganisms. We have shown that certain ecosystem functions are positively
correlated with microbial biodiversity. However, attempts to understand the
mechanism underlying that relationship have borne little fruit. We argue that
to overcome this challenge we should look for inspiration from population
genetics and the process of genotype-phenotype mapping. In particular, we should look
more agnostically at microbial
communities to identify taxa that regulate ecosystem-scale functions. In
addition, we should attempt to understand the complex covariance that exists
between microbial communities, biogeographic patterns, and the underlying abiotic
environment. Finally, we reframe the question "Is microbial biodiversity
related to ecosystem function?" as "What is the unique contribution of the
microbiome to ecosystem function independent of the environment?"

### There is evidence that microbial diversity matters

There is evidence that microbial communities should matter for regulating
ecosystem function. Some studies have manipulated the connection between
environmental factors and microbial community composition through common garden
or reciprocal transplant experiments. These studies frequently report different
rates of ecosystem functions for different microbial communities under the same
environmental conditions [@reed2007]. This has been observed for plant
decomposition [@glassman2018, @strickland2009], plant phenology
[@panke-buisse2015], soil nitrogen cycling [@balser2005], and soil greenhouse
gas emissions [@cavigelli2000], among other functions. Therefore, variation in microbial
community composition can be associated with variation in ecosystem function.

Studies using a comparative approach rather than a manipulative one have also
demonstrated a link between microbial biodiversity and ecosystem function.  In
general, there is a significant correlation between belowground biodiversity
and multiple ecosystem functions [@jing2015, @delgado-baquerizo2016,
@delgado-baquerizo2017, @wagg2014]. It is unclear whether the positive
diversity-function relationship is due to the independent effect of variation
in community composition on ecosystem function or due to environmental
conditions that select for both diverse communities and high rates of ecosystem
function. Determining why we see this diversity-function relationship depends
on identifying the components of the community that are related to ecosystem
function.

### We need a new perspective on microbial biodiversity-ecosystem function relationships

Most sudies of microbial community function focus on either functional gene
abundance or community diversity.
In the case of functional genes, qPCR or shotgun metagenomic sequencing is used to
estimate the abundance of a gene or transcript that is a putative marker for a
microbial process (and thus a marker for the functional group that performs
that process). For example, the gene _mcrA_, which encodes a subunit of the
enzyme that performs the final step in methanogenesis, is commonly used as a
marker for methanogenesis and for the methanogen functional group. Other
examples include _pmoA_ for methanotrophy, _nifH_ for nitrification, and _nosZ_ for
denitrification. The abundance of these markers is often hypothesized to
predict the rate of their associated processes (for example, the
abundance of _mcrA_ is expected to relate to the rate of
methanogenesis). There are examples where this relationship is present
[@freitag2009, @freitag2010, @schnyder2018].  However, a review of such studies
found that the abundance of a functional gene or transcript is rarely
correlated with the rate of the corresponding process and depends on the
function of interest, with most effects either negative or not significant
[@rocca2015].

The second aspect of microbial community structure hypothesized to predict
ecosystem function is taxonomic or functional diversity. Diversity is either
estimated from sequence variants of a barcode gene such as the 16S rRNA gene or
manipulated through some proxy of diversity such as sequential dilution or
varying filter sizes. Studies that experimentally manipulate diversity
generally find a relationship between the applied diversity treatment and
ecosystem function, including methanotrophy [@schnyder2018], phosphorus
leaching [@wagg2014], greenhouse gas emissions [@wagg2014], decomposition
[@maron2018], and nitrogen cycling [@philippot2013]. However, in many cases
richness is confounded with other factors such as the presence or absence of
major phylogenetic groups (nematodes, fungi) or abundance of microbial cells.
In addition, assembled microbial communities risk confounding diversity with
composition, for example by more likely including the most productive taxa in
the highest diversity treatment. In general, the relationship between microbial
diversity is positive and significant, but adds little variance over abiotic
conditions or macroorganismal diversity [@graham2016, @jing2015].

Despite the positive diversity-function relationship for microbes, models that
incorporate microbial features only marginally improve predictions over
environmental variables [@graham2016]. In addition, we've made little progress
in identifyng the components of microbial communities that drive these
diversity-function relationships [@rocca2015].
Overall, functional gene abundance and community diversity improve models of
ecosystem function less than one third of the time and increase variance
explained by an average of only 8 percentage points [@graham2016].

Up to this point, microbial biodiversity-ecosystem function research has
followed a similar path to macrobial biodiversity-ecosystem function research,
particularly compared to plant communities. However, there are important
differences between plants and microbes that might inform how we pursue this
research going forward. For example, microbial communities are more diverse and
contain greater numbers of individuals than plant communities and are thus more
difficult to sample comprehensively. Unlike plants, functional groups of
microbes are rarely determined through observation or direct measurement, but
rather must be indirectly inferred based on taxonomic assignment of marker
genes (such as 16S rRNA genes), by sequencing functional genes from
metagenomes, or by quantifying specific functional genes using qPCR.  This adds
uncertainty in our ability to estimate the size of functional groups.  The
ability to experimentally manipulate diversity or community composition is much
more limited for microbes. Given all of these differences (and many others),
the challenge of linking microbial communities to ecosystem functions may be
less like that of plant communities and more analogous to the complex task of
linking genomic variation to organismal phenotypes. For this reason, we propose
that integrating traditional biodiversity-ecosystem function research with
ideas from genotype-phenotype mapping could provide the new perspective our
discipline needs.

### Genotype-phenotype as a source of inspiration

Given the relative lack of success to date, it is time to rethink how we
approach the challenge of mapping microbial community structure to ecosystem
function. Microbial ecologists are not the only biologists who are attempting
to determine the relationship between a complex set of highly-variable data and
an aggregate function. This kind of “many-to-one” mapping is analogous to the
challenge of identifying the genetic basis of complex traits in organismal
populations. In such genotype-phenotype mapping studies, a population
exhibits variation in a phenotype (e.g., height or disease state) as well as
variation in thousands of genetic loci or single nucleotide polymorphisms (SNPs).
To identify the genetic basis for a trait, investigators sample from this
population and correlate phenotype with genotype. In most cases, there are many
more loci than individuals and we do not know whether the SNPs are causally
linked or are simply in linkage disequilibrium with a causal mutation.

There are a number of parallels between the challenge faced by organismal
biologists and that facing microbial community ecologists. They are both
“many-to-one” mapping challenges which involve large numbers of statistical
comparisons. Both are attempting to identify causal relationships that are
potentially confounded by complex patterns of covariation. For example,
with local selection or distance-decay. Further, a positive association may
either indicate a causal relationship or simply an association between a
significant taxon and a causal taxon (i.e., microbial linkage
disequilibrium). There is often no strong expectation about which entities
(i.e., which genomic regions or which microbial taxa) are most likely to be
causally-related to phenotype or function, and thus “agnostic” approaches are
needed.  Both are sensitive to exactly how the “mapping” question is asked, and
both ultimately require manipulation (of genes or taxa) to establish causation. We
describe each of these parallels below in more detail and provide an example
of how these ideas could be applied to microbial data.

### The importance of a taxonomically “agnostic” approach

[This is super redundant with the previous section.]

The traditional approach to macroorganismal biodiversity-ecosystem function
research is to measure or manipulate the diversity of a taxonomic group (e.g.,
plants) and look for an association with the function performed by that group
(e.g., primary productivity). In the broadest sense, we can think of plants as
a “functional group,” i.e., a group of taxa united by their ability to perform
a particular ecosystem function. Ecologists may further divide a functional
group (e.g., plants) into smaller functional groupings (e.g., forbs), defined
by the details of how they perform their particular ecosystem function.

However, for microbes our knowledge of functional groups is much more limited.
From a very limited number of cultured isolates we have a provisional
understanding of which microbes might be involved in some ecosystem functions.
And by sequencing the genomes of these isolates, we have identified genetic
markers for certain functions. But most microbial taxa remain uncultured and we
do not know the function of most microbial taxa detected in environmental
samples [@hug2016, @martiny2019, @steen2019]. In addition, there have been
recent discoveries of functions in unexpected taxonomic groups, for example
methanogenesis by fungi and cyanobacteria, a function previously considered
restricted to the Archaea [@bizic-ionescu2019, @lenhart2012]. Therefore,
estimating microbial functional group abundance can be challenging.

Beyond, for example, identifying previously undocumented methanotrophs that are
related to methane oxidation, it's possible that more complex ecological
regulatory patterns exist in microbial communities. In other words, it may
be that methanotroph abundance does not limit methanotrophy in a given ecosystem.
[Something about metabolic networks...]
Instead, the rate
of methane flux to the atmosphere could be controlled by methanotrophs,
methanogens, bacteria that produce the fermentative byproducts that methanogens
depend on, nutrient cyclers that regulate the availability of nitrogen or
phosporus, or even grazers and other predators or pathogens. In other words, the regulation of
ecosystem function by microbial communities could be direct or indirect, much
like genetic regulation within an organism.

Because of this, it would be prudent to look more agnostically at microbial
communities to identify taxa or groups of taxa that are important for
predicting the rate of ecosystem functions rather than assuming that the
genetic markers we have provisionally identified for a given function represent
the most likely taxa involved. This agnostic approach is analogous to the
approach of many genotype-phenotype mapping studies (e.g., genome-wide
association studies or GWAS), which often look for associations between a
phenotype and loci anywhere in a genome.

### Controlling for population stratification

It is widely accepted that organisms, including microorganisms, exhibit
population stratification due to geographic and environmental separation
[@martiny2006, @wright1943]. Genome-wide association studies generally control
for population stratification using shared ancestry among cases and controls when
modeling the connection between genotype and phenotype. The classic example is
the latitudinal gradient of both height and genotypic similarity in Europe,
which results in spurious associations between human height and genetic
variation [@berg2019, @novembre2008, @sohail2019]. Perhaps a more illustrative
example is the correlation between chopstick usage and certain genetic markers
in Asian populations, which is unlikely to be due to an actual genetic basis
for chopstick skills. Spurious associations arise
because of heterogeneity in population structure where any genetic variant in a
population with an extreme phenotype will tend to be associated with that
phenotype whether it is related or not.

Genome-wide association studies correct for stratification using a variety of methods.
Generally, they ignore the underlying environmental and spatial
distance between samples and instead use shared ancestry as a proxy for local
selection and assortative mating.  A common approach is to perform a regression
of phenotype and shared ancestry (computed as the first one or more principal
components of the genotype matrix) and then use the residuals from this model
as the values for phenotype in a
subsequent regression using the genotypes directly [@price2006]. This principal
component analysis is designed to test the effect of individual genes after
removing shared ancestry among individuals. Another approach, employed in our
example, is variance component modeling (or mixed modeling, heirarchical
modeling, etc.) where genotypic similarity is included as a covariate in the
model [@kang2010].

While these represent two common statistical approaches to
control for population stratification, it is also considered good practice to
control for stratification experimentally. This can be accomplished by 
selectng a study population that is relatively homogeneous. Otherwise, even
large, well-replicated GWAS studies can result in
false positives. For example, it was recently revealed that GWAS studies on
height of Europeans failed to adequately control for stratification. When their
results were confirmed using a similarly sized, but more homogeneous study
population, many of the genetic markers for height were no longer significant [@berg2019,
@sohail2019].

[This paragraph needs work]

Typically, microbial biodiversity-ecosystem function studies do not account for
population stratification (i.e., community similarity across samples), although
there are some exceptions [@meyer2019, @qin2012, @lloyd-price2019].  However,
community similarity (the community analogue of shared ancestry among
organisms) is not as tightly linked to geography or environment as is shared
ancestry among organisms, and it could be very useful to account for these
separately in microbial studies, especially if one is particularly interested
in how composition alters function independent of the underlying environmental
variation.

### What might this look like/More about the analogy

The relationship between structure and function could be complex and indirect.
i.e. not just methanogens making methane, but metabolic and regulatory networks

A significant correlation between a taxon and a function could mean a number of
things.
1. That microbe is performing that function right now and its abundance is
   related to function
2.

This whole compositional data/relative abundance thing also seems problematic
for correlating abundance with function. Really what we're saying is "this read is differentially abundant in low vs. high flux ecosystems." Not abundance is _correlated_ with flux.
1. It's not **abundance**.
2. An increase in relative abundance in one taxon implies nothing about the
   abundance of that taxon

   wootincrease in relative abundance in one taxon implies nothing about the
abun
3. It is perhaps somewhat more likely that