### 2019-11-12 manuscript/manuscript_rev1.Rmd
Incorporating Brendan's edits

---
title: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
output: 
    word_document:
        reference_docx: reference.docx
bibliography: library.bib
csl: philosophical-transactions-of-the-royal-society-b.csl 
---

```{r setup, include=FALSE}

library(tidyverse)
library(knitr)
library(flextable)
opts_chunk$set(echo = FALSE)

```

# Abstract

Microbial physiological processes are intimately involved in nutrient cycling.
However, it remains unclear to what extent microbial diversity or community
composition is important for determining the rates of ecosystem-scale
functions. There are numerous examples where microbial diversity is correlated
with the rate of an ecosystem function, but exactly how microbial communities
“map” onto ecosystem functions remains unresolved.  This uncertainty limits our
ability to predict and manage crucial microbially mediated processes such as
nutrient loss and greenhouse gas emissions. Our lack of understanding stems
from the relatively large diversity of microorganisms, the difficulty in
directly identifying functional groups, and our limited ability to manipulate
microbial community attributes. For this reason, we propose that integrating
traditional biodiversity-ecosystem function research with ideas from
genotype-phenotype mapping could provide the new perspective our discipline
needs. We identify two insights from genotype-phenotype mapping that could be
useful for microbial biodiversity-ecosystem function studies: the concept of
“agnostic” mapping and controlling for population stratification to identify
microorganisms that uniquely contribute to ecosystem function. We illustrate
the potential for these approaches to elucidate microbial
biodiversity-ecosystem function relationships by analyzing a subset of
published data measuring methane oxidation rates from tropical soils. We assert
that combining the approaches of traditional biodiversity-ecosystem function
research with ideas from genotype-phenotype mapping will not only generate
novel hypotheses about how complex microbial communities drive ecosystem
function, but also help scientists predict and manage changes to ecosystem
functions resulting from human activities.

## Introduction

Ecology is broadly focused on understanding biodiversity and how that
biodiversity shapes the ecosystems that humans depend on. Many ecosystem
processes essential to all of life are mediated by microorganisms and therefore
understanding the relationship between microbial biodiversity and ecosystem
function is important [@schimel1998, @singh2010]. Certain ecosystem functions
are correlated with microbial diversity, indicating that we should be able to
determine what aspects of microbial biodiversity influence ecosystem function.
However, attempts to describe that mapping have borne little fruit [@rocca2015,
@graham2016]. We argue that to overcome this challenge we should look to other
successful attempts at mapping biological variation onto higher order
processes. In particular, population genetics and the process of
genotype-phenotype mapping provide a number of potentially useful insights. For
example, genotype-phenotype mapping often makes few assumptions about the
nature of the map, i.e., it is “agnostic.” In addition, population geneticists
have developed rigorous methods for reducing potentially confounding
relationships such as geographic structuring of populations. Finally, the
ultimate goal of genotype-phenotype mapping is to identify the unique
contribution of genotype to phenotype separate from other drivers of phenotypic
variation such as the environment. Inspired by this, we suggest a reframe of
the question, "is microbial biodiversity related to ecosystem function?" to, "what
is the unique contribution of the microbiome to ecosystem function independent
of the environment?"

### There is evidence that microbial biodiversity matters for ecosystem function

There is increasing evidence for microbial biodiversity-ecosystem function
relationships. For example, there are positive correlations between microbial
diversity and ecosystem multifunctionality for a variety of ecosystems and most
major lineages of microorganisms [@jing2015, @delgado-baquerizo2016,
@delgado-baquerizo2017]. Adding microbial diversity or abundance to ecosystem
models can in some cases improve model accuracy [@graham2016]. Contrived
communities that vary in richness and communities created through sequential
dilution or varying filter size to generate differences in diversity can also
exhibit differences in ecosystem function [@philippot2013, @wagg2014, @schnyder2018,
@maron2018].  Finally, reciprocal transplant and common garden experiments that
manipulate the connection between community composition and environment reveal
differences in ecosystem function for communities of distinct
origins [@cavigelli2000, @strickland2009, @glassman2018]. Given these
relationships, many investigators have now moved on to the challenge of
identifying the aspects of microbial biodiversity (e.g., specific taxa, genes,
functional groups, etc.) that influence a given ecosystem function; however,
this has proven especially challenging.

### The mapping between microbial biodiversity and ecosystem function has been elusive

Most studies that attempt to identify the aspects of microbial biodiversity
that influence a given ecosystem function focus on “functional” gene or
transcript abundance.  In this case, qPCR or shotgun metagenomic sequencing is
used to estimate the abundance of a gene or transcript that is a putative
marker for a microbial process (and thus a marker for the functional group that
performs this process). For example, the gene _mcrA_, which encodes a subunit of
the enzyme that performs the final step in methanogenesis, is commonly used as
a marker for methanogenesis and for the methanogen functional group. Other
examples include _pmoA_ and methanotrophy, _nifH_ and nitrification, and _nosZ_ and
denitrification. The abundance of these markers is often hypothesized to be
predictive of the rate of their associated processes (for example, the
abundance of _mcrA_ is often hypothesized to be related to the rate of
methanogenesis).

Some ecosystem functions in certain ecosystems can be predicted from the
abundance or transcriptional activity of genetic markers for those functions.
For example, soil methane production and consumption can under some
circumstances be predicted from the genetic markers _mcrA_ and _pmoA_
[@freitag2009, @freitag2010, @mccalley2014]. However, for most ecosystem
functions the abundance of a functional gene or transcript is rarely positively
correlated with the rate of the corresponding process [@rocca2015]. The
cases where there is a positive correlation tend to be restricted to
agricultural ecosystems and for certain functions within the nitrogen cycle
[@rocca2015]. In general, including aspects of microbial biodiversity
(e.g., functional gene abundance or diversity) improve models of ecosystem
function less than one third of the time, and increase variance explained by an
average of only 8 percentage points over environmental variables [@graham2016].

## Genotype-phenotype mapping as a source of inspiration

In the approaches described above, microbial ecologists often use microbiome
data to infer taxonomic composition, essentially creating species lists from
data such as 16S rRNA marker genes or shotgun metagenomes. Interpreting
microbiome data in this way has allowed us to use approaches from
biodiversity-ecosystem function research (which are often focused on taxonomic
or functional groups), but it has generally not been useful for creating more
detailed descriptions of the relationship between microbial biodiversity and
ecosystem function. But this approach is not the only way one could determine
the relationship between a complex set of highly-variable data and an aggregate
function.

This kind of “many-to-one” mapping is analogous to the challenge of identifying
the genetic basis of complex traits in organismal populations. In such
“genotype-phenotype” mapping studies, a population exhibits variation in a
phenotype (e.g., height or disease state) as well as variation in potentially
thousands of single nucleotide polymorphisms (SNPs). To identify the genetic
basis for a trait, investigators sample from this population and correlate
phenotype with genotype. While some phenotypes (e.g., the propensity for
diseases such as Parkinson’s) are controlled by a single genetic locus
[@kerem1989, @macdonald1992], most traits depend on a large number of genes
that control variation in that phenotype [@hindorff2009, reich2001]. In
addition, there is often no a priori expectation about which regions of the
genome control that trait so we must search for genetic markers
throughout the genome. If a marker is significantly correlated with the
phenotype of interest, this either indicates it is inside of a gene with a
direct or indirect effect on phenotype or that it is in linkage disequilibrium
(i.e., non-random association between two alleles) with a causal gene.

There are a number of parallels between this challenge faced by organismal
biologists and that facing microbial community ecologists. They both involve
large numbers of statistical comparisons. Both are attempting to identify
causal relationships that are potentially confounded by complex patterns of
covariation. There is often no strong expectations about which entities (i.e.,
which genomic regions or which microbial genes or lineages) are most likely to
be causally-related to phenotype or function, and thus “agnostic” approaches
are needed. For some ecosystem functions, it is possible that a single
microbial taxon could substantially influence its rate. For example, methane
flux from permafrost in Sweden may be controlled by a single taxon [@mccalley2014]. But
for most ecosystem functions, there could be many taxa of small effect that
contribute to the rate of ecosystem function. Finally, both ultimately require
manipulation (of genes or taxa) to establish causation.

### The importance of a taxonomically “agnostic” approach

Most microbial biodiversity-ecosystem function research up to this point has
used an approach analogous to that used by plant ecologists studying
biodiversity-ecosystem function relationships. This approach is to measure or
manipulate the diversity of a taxonomic group (e.g., plants) and look for an
association with the function performed by that group (e.g., primary
productivity). We can think of plants as a “functional group,” i.e., a group of
taxa united by their ability to perform a particular ecosystem function. For
microbes, estimating functional group abundance can be much more challenging.
From a small number of cultured isolates, we have a provisional understanding
of which microbes might be involved in some ecosystem functions. By sequencing
the genomes of these isolates, we have identified genetic markers for certain
functions, which we call “functional genes.” But most microbial taxa remain
uncultured and we do not know the function of most microbial taxa detected in
environmental samples [@hug2016, @martiny2019]. In addition, there have
been recent discoveries of functions in unexpected taxonomic groups, for
example methanogenesis by fungi and cyanobacteria, a function previously
considered restricted to the Archaea [@bizic-ionescu2019, @lenhart2012].

As illustrated earlier, these functional markers are not correlated with their
corresponding ecosystem function in most ecosystems and for most processes. In
addition, they provide little explanatory power to models of ecosystem
function. Because of this, it might be prudent to look more agnostically at
microbial communities to identify taxa, groups of taxa or genes that are
important for predicting the rates of ecosystem functions rather than assuming
that the genetic markers we have provisionally identified for a given function
represent the most likely taxa or genes involved. This agnostic approach is
analogous to the approach of many genotype-phenotype mapping studies (e.g.,
genome-wide association studies), which often look for associations between a
phenotype and loci anywhere in the genome.

Beyond finding new physiologies in unexpected places, there are other reasons
for looking agnostically. In the case of microbial functions, it may be that
the organism that performs a function is not the limiting factor for the rate
of that function. For example, the rate of soil-to-atmosphere methane flux
could be limited by methanogens or methanotrophs or the balance of the two. It
could also be limited by the bacteria that produce the fermentative byproducts
that methanogens depend on. Or there could be indirect limitation by organisms
that liberate nitrogen or phosphorus into mineral forms. In other words, the
influence of microbial communities on the rate ecosystem function could
represent a complex metabolic network much like the regulation of gene
expression in organisms that partially determine their phenotype. These broader
patterns of biodiversity-ecosystem function relationships would be invisible to
any study that focuses on putative functional groups without considering the
reads that underlie those functional groups.

### Controlling for population stratification

It is widely accepted that organisms, including microorganisms, exhibit
population stratification due to geographic and environmental separation
[@wright1943, @martiny2006]. This can lead to spurious associations between
phenotypes and genetic markers that are at high frequency in isolated
sub-populations. Association studies generally control for population
stratification by accounting for shared ancestry among organisms in a
population when modeling the connection between genotype and phenotype.
Typically, microbial biodiversity-ecosystem function studies do not account for
population stratification (i.e., community similarity among ecosystems),
although there are some exceptions [@meyer2019, @qin2012, @lloyd-price2019].
Community similarity (the community analogue of shared ancestry among
organisms) is not as tightly linked to geography or environment as is shared
ancestry, and it could be very useful to account for these separately in
microbial studies, especially if one is particularly interested in how
composition alters function independent of the underlying environmental
variation.

Genome-wide association studies correct for stratification using a variety of
methods. Generally, they ignore the underlying environmental and spatial
distance between samples and instead use shared ancestry as a proxy for local
selection and assortative mating. A common approach is to perform a regression
of phenotype and shared ancestry (computed as the first one or more principal
components of the genotype matrix) and then use the residuals from this model
as the values for phenotype in a subsequent regression using the genotypes
directly [@price2006]. This principal component correction is designed to test the
effect of individual genes after removing the effect of shared ancestry among
individuals. Another approach, employed in our example, is variance component
modeling (or mixed modeling, hierarchical modeling, etc.) where genotypic
similarity is included as a covariate in the model to control for
stratification while testing the genotype-phenotype connection [@kang2010].

If we control for covariates such as community similarity, geographic distance,
or environmental similarity, it changes the nature of our question. For
example, if we test the correlation between the relative abundance of a taxon
and the rate of methane flux we are asking “is this taxon correlated with
methane flux?” If we find a significant result, that may be because variation
in the abundance of that organism directly or indirectly contributes to methane
flux. However, it might also be that that organism lives only in ecosystems
that happen to have a high rate of ecosystem function. We would not be able to
distinguish these outcomes with this simple correlation. If we then include
environmental variables or environmental similarity as a covariate, we could
start to ask, “Is the taxon uniquely associated with function independent of
the environment?” By “uniquely associated”, we mean those taxa associated with
function irrespective of environmental conditions, local community structure,
or spatial proximity. This is not only a more specific question than is usually
asked in microbial biodiversity-ecosystem function studies, but it is also a
more appropriate one, especially if one is interested in how to incorporate
microbial community data into ecosystem models. Finally, it is interesting on
its own to understand whether microbial communities are selected by the
underlying environmental conditions to produce a particular rate of ecosystem
function (the conduit) or whether there is variation in community composition
independent of the environment that explains some variation in ecosystem
function.

## An example: high-affinity methane oxidation

To illustrate the ideas outlined above, we reanalyzed a subset of previously
published data from a paper that has demonstrated one successful approach
for applying genotype-phenotype mapping to microbial communities [@meyer2019]. In our
reanalysis, we do not intend to challenge the conclusions from that paper, but
instead we want to demonstrate how to perform this type of study for microbial
ecologists unfamiliar with association studies. A full description of the study
design, samples, and data generation can be found in that manuscript. Briefly,
these data were gathered from intact soil cores taken from diverse ecosystems
of the Congo Basin in Gabon, Africa. Cores were incubated in the laboratory
under different concentrations of methane to identify the rates of specific
methane cycling pathways. For this example, we will analyze data from just one
of these pathways, high-affinity methane oxidation (the oxidation of
atmospheric concentrations of methane), which we will refer to simply as
“methane oxidation”. In addition, we only include amplicon sequences from the
DNA-inferred community and not the RNA-inferred community, both of which are
presented in the original paper (@meyer2019). The data we analyzed include methane
oxidation rates, amplicon sequence variants (ASVs; @callahan2017) generated using DADA2
and inferred from 16S rRNA gene sequences, _pmoA_ abundance estimates (via qPCR),
latitude and longitude, and four environmental covariates (soil moisture, bulk
density, carbon, and nitrogen).

Analyses were conducted in the R statistical environment using the phyloseq
package [@mcmurdie2013, @rcoreteam2019]. The relative abundances of ASVs were corrected using the variance
stabilizing transformation from DESeq2 [@love2014,
@mcmurdie2014]. We first test typical measures of
microbial structure including functional gene abundance and community richness,
which was estimated using the breakaway package [@willis2015]. We then demonstrate
significant covariation between community structure (estimated as Bray-Curtis
distance using vegan), environmental variation (euclidean distance), and
geographic distance (euclidean distance) using Mantel tests [@bray1957, @oksanen2019]. Finally, we
present one approach to identifying taxa which are significantly associated
with function independent of the environment by fitting variance component
models using varComp to test the relationship between relative abundance of
each ASV and methane oxidation rate [@kang2010, @qu2013]. To illustrate how
including different
covariates (environmental, geographic, and community) can result in different
conclusions about which taxa are associated with function, we fit this model
with and without random effects variance components for environmental
similarity, geographic site ID, and Bray-Curtis similarity. Significant taxa
were determined by controlling the false discovery rate at q-value < 0.05 [@storey2002].
Figures were created using ggplot2 [@wickham2019].

### Results & Discussion

Microbial biodiversity-ecosystem function studies typically test functional
group abundance or community richness as it relates to ecosystem function. In
our case, methane oxidation rate was not significantly correlated with the pmoA
gene abundance or 16S read richness (Table 1, Figure 1). To demonstrate that
the covariance structure of the data might alter our conclusions about which
taxa regulate ecosystem function, we tested collinearity between each pair of
distance matrices for community, environment, and geography. We found a
moderate and significant correlation between community composition and
environmental variation, geography and community composition, and geography and
environmental variation (Table 2, Figure 2). To visualize this population
stratification, principal coordinate plots show that beta diversity of samples
separated by site ID and by ecosystem type (wetland or upland; Figure 2), which
indicates substantial spatial and environmental structuring of microbial
populations. This means that the presence or abundance of certain genetic
markers (taxa) will be elevated in specific ecosystems. In this case,
high-affinity methane oxidation is typically greater in upland ecosystems than
in wetland ecosystems and so any taxa differentially abundant in uplands will
be correlated with methane oxidation regardless of their involvement in that
process. It is necessary to control for this structuring to rigorously identify
associations between taxa and function.

To demonstrate this approach, we tested the effect of the relative abundance of
each ASV on methane oxidation rate while controlling for different sets of
covariates including environment, geography, and community. After controlling
the false discovery rate, 460 unique 16S reads were identified as significantly
correlated with function when no covariates were included in the model. We
found different numbers of taxa significantly associated with methane oxidation
depending on which covariates were included in the model (Table 3). Each of
these sets of taxa represent different versions of the biodiversity-ecosystem
function mapping question. For example, controlling for environmental variation
tests for taxa that are related to function independent of environmental
conditions. Similarly, controlling for geographic distance tests for taxa that
are related to function independent of geographic proximity; geographic
proximity could influence function via unmeasured environmental variation that
is spatially structured or differences in dispersal history among sites due to
physical distance. In the model that controlled for all three covariates
(community, environment, and geography), only six ASVs were significantly
correlated with methane oxidation rate (Figure 3). These taxa could be useful
indicators for methane oxidation rate across space and different ecosystems. In
addition, researchers could manipulate their abundance in order to better
understand the biological drivers of variation in methane oxidation.

Notably, these six taxa fall into three genera and one class with cultured
representatives that are not known to consume methane [@belova2009, @chang2011,
@domeignoz-horta2019, @fritz2004]. These taxa could be
related to ecosystem function in multiple ways. The most interesting
possibility is that each of these taxa is statistically related because it is
causally connected to the function. This could be direct, for example an
organism that consumes methane, or indirect, for example an organism that
regulates substrates necessary for methane oxidizers. Alternatively, a
significant association could occur for non-causal reasons. For example, any
organism that tends to be in high abundance where methane oxidation rates are
high would be correlated with methane oxidation, even if it has no causal
relationship. This could be because such an organism is favored under the same
conditions that favor methane oxidation. Such covariation can
drive associations that are not causal, but the effects of covariation would
have been reduced by controlling for covariates in our tests.

This approach is powerful because it accounts for the covariances that arise
from the fundamental ecological processes that drive community assembly and
underlie biogeographical patterns [@martiny2006, @vellend2010]. It is also
powerful because it is
focused on a much more specific version of the “biodiversity-ecosystem
function” question, i.e., it asks “which taxa are uniquely associated with
function?” By “uniquely associated,” we mean those taxa associated with
function irrespective of environmental conditions, local community structure,
or spatial proximity. This is not only a more specific question than is usually
asked in microbial biodiversity-ecosystem function studies, but it is also a
more appropriate one, especially if one is interested in how to incorporate
microbial community data into ecosystem models. These models depend on
identifying taxa that add explanatory power beyond that provided by other
factors, such as environmental conditions. Earth system models would benefit
most from microbial bioindicators that capture variation unexplained by other,
more easily measured variables (such as certain environmental conditions). The
approach we describe is an important step toward identifying such biomarkers.

## Caveats and future directions

Once taxa have been identified with an association test (such as the one we
outline above), there are multiple ways they could be used for future study.
One approach common in genetics, particularly for markers of genetic disorders,
is to generate a polygenic score based on the summed effect of many genes on a
phenotype of interest, such as the probability of developing a disorder. A
similar aggregate bioindicator could be generated for ecosystems that would
summarize the probability of having a particular ecosystem function or rate.
This would be accomplished by measuring the abundance of the taxa identified in
an association study and determining their association with the rate of an
ecosystem function. Alternatively, the identified taxa could be incorporated
into a structural equation model in an attempt to better understand the
individual effects and interactions among taxa as they contribute to the rate
of ecosystem function. This might give an indication of the relative importance
of different taxa as compared to other factors, such as environmental
variables.

Ultimately, the relationships identified in any comparative mapping study
must be verified. For genotype-phenotype studies in organisms there are
multiple ways that this verification is
accomplished. In some cases, organisms can be artificially selected for a
particular phenotype (e.g., through experimental evolution in an environment
that favors the phenotype of interest) and the genetic changes that occur in
response to selection can be compared to those identified via mapping studies
(such as GWAS). An analogous approach for microbial biodiversity-ecosystem
function studies would be to apply artificial ecosystem selection (sensu @swenson2000)
on a given function and compare the taxa that change in response to selection
with those identified via a comparative approach (such as the one illustrated
in our example).

The most common way that loci identified in a genotype-phenotype mapping study are verified is
through manipulative genetics. The identified loci can be knocked out or
over-expressed and the effect on phenotype compared to that predicted from
mapping studies. In the case of microbial biodiversity studies, it may be
possible to inhibit a particular functional group through the use of a specific
antimicrobial or a chemical inhibitor or by using phages with high
host-specificity [@maxson2016], but this is not generally possible. In some cases, we
may be able to isolate a microorganism of interest in pure culture and add it
back to an ecosystem, transiently increasing its abundance (roughly analogous
to “overexpressing” a gene). A greater focus on culture-based approaches could
increase the success of these kinds of microbial enrichments. Finally,
synthetic communities (contrived assemblages of microorganisms) may be the most
powerful way to test hypotheses about microbial biodiversity-ecosystem function
relationships, but currently these approaches are limited by the small number
of taxa that can be routinely cultured from most environments.

There are a number of limitations to the biodiversity-ecosystem function
mapping approach we describe, some in common with organismal mapping studies
and others unique. For example,
simple linear models such as the variance component model used in this study
are typical for genetics studies, but may not be the best way to identify
correlations for microbiomes due to the unique challenges of microbial data.
These are zero-inflated compositional data with differences in sequencing
depth across samples. We have addressed these challenges by using a variance
stabilizing transformation, but other models that test differential
abundance and differential variance that can control for differences in
sequencing depth and are robust to zero-inflation might be more appropriate
(e.g., CORNCOB @martin2019).  Alternatively, clustering reads at higher
taxonomic levels could provide a more continuous variation in taxon abundances
across ecosystems by reducing the number of “bins.” The tradeoff for this is
that it introduces bias based on the completeness of our taxonomic databases
and the accuracy of 16S-based taxonomic assignment and it removes reads that
lack a taxonomic assignment. Here, we chose to test ASVs at the level of the
individual read so as not to bias our results by losing reads that lack a
taxonomic assignment at a particular level.

Experimentally, future studies could improve on our example by sampling a more
homogeneous set of ecosystems. Our survey includes grasslands, pastures,
forests, peatlands, and mineral soil wetlands among others. These ecosystems
represent a range of moisture conditions that could regulate the abundance and
activity of methane oxidizers and access to methane and oxygen, which methane
oxidizers rely on. While this captured substantial variation in methane
oxidation rates, sampling from such diverse ecosystems could result in spurious
associations between taxa and function. For example, taxa differentially
abundant in upland ecosystems that are unrelated to methane oxidation might
appear correlated simply due to their presence in those ecosystems with high
oxidation rates. Future studies could try restricting their search to a more
homogeneous population of ecosystems specific to the question at hand.

## Conclusion

Microbial biodiversity-ecosystem function research has demonstrated positive
correlations between diversity and ecosystem function. However, the abundances
of microbial functional groups (as currently defined) are often not good
predictors of ecosystem function and commonly do not add substantial
explanatory power to ecosystem models. Therefore, a new perspective on how to
determine the relationship between microbial communities and ecosystem
functions is sorely needed. Organismal biologists have over a hundred years of
experience identifying relationships between complex sets of highly-variable
data (genotypes or genome sequences) and aggregate functions (organismal
phenotypes). We assert that combining the approaches of traditional
biodiversity-ecosystem function research with ideas from genotype-phenotype
mapping could provide this new perspective. This integration could not only
make underutilized approaches such as covariate modeling and artificial
selection more available to microbial ecologists, but also provide instructive
examples of how best to conceive of microbial biodiversity-ecosystem function
questions. If this integration is successful, it is possible that in the
not-so-distant future our field will be able to robustly identify taxa, genes,
or even molecules that will allow us to accurately predict the response of
ecosystems to environmental change. Doing so will not only generate novel
hypotheses about how complex microbial communities drive ecosystem function,
but also help scientists predict and manage changes to ecosystem functions
resulting from human activities.

# Acknowledgments

This project was supported by the National Science Foundation – Dimensions of
Biodiversity program (DEB 14422214), the National Science Foundation Graduate
Research Fellowship Program (DGE 1255832), and the ARCS Foundation Florence and
Mike Nudelman Scholarship. The data re-analyzed here were generated with
permission of the Government of Gabon, Centre National de la Recherche
Scientifique et Technologique (Permit No AR0035/14/MESRS/CENAREST/CG/CST/CSAR)
and the assistance of F. Bivigou, H. Memiaghe, L. Tchignoumba, E.  Tobi, and I.
Akendengue. The University of Oregon, the Gabon-Oregon Transnational Center on
Environment and Development, the Smithsonian Conservation Biology Institute,
and Shell Gabon provided financial and logistical support for the original
study re-analyzed here. We also thank W. Cresko for insightful discussions regarding
population genetics and H. Tavalire for contributions to the statistical
analyses.

# References

::: {#refs} 
:::

# Tables and Figures

```{r copy_num}
model_out <- read.csv('../output/model_out.csv')
flextable(model_out) %>% 
  compose(i = 1, j = 1, 
    value = as_paragraph(as_i("pmoA"), " copy number")) %>% 
  compose(part = "header", j = 4, 
    value = as_paragraph("t-statistic")) %>% 
  compose(part = "header", j = 5, 
    value = as_paragraph("p-value")) %>% 
  autofit()
```

Table 1. Functional gene abundance and ASV richness are not good predictors of
methane oxidation rate. Linear models predicting methane oxidation rate from
measures of microbial community structure.

```{r mantel}
mantel_table <- read.table('../output/mantel_table.tsv')
flextable(mantel_table) %>% 
  compose(part = "header", j = 2, 
    value = as_paragraph("Mantel statistics (r)")) %>% 
  compose(part = "header", j = 3, 
    value = as_paragraph("95% upper quantile of permutations")) %>% 
  compose(part = "header", j = 4, 
    value = as_paragraph("p-value")) %>% 
  autofit()
```

Table 2. Mantel tests for each pair of dissimilarity matrices. Community
distance matrix was based on Bray-Curtis distance while both environment and
geography distance matrices were based on Euclidean distance. P-values
determined by permutation test with 999 permutations.

```{r n_taxa}
n_taxa <- read.table('../output/comp_adds_removes.tsv')
flextable(n_taxa) %>% 
  compose(part = "header", j = 1, 
    value = as_paragraph("Terms")) %>% 
  compose(part = "header", j = 2, 
    value = as_paragraph("Removed")) %>% 
  compose(part = "header", j = 3, 
    value = as_paragraph("Added")) %>% 
  compose(part = "header", j = 4, 
    value = as_paragraph("Significant")) %>% 
  autofit()
```

Table 3. Number of significant taxa after including each set of covariates in a
variance component model. Removed and Added columns are relative to the
no-covariate model. Significance is determined by controlling the false
discovery rate at q-values < 0.05.

```{r methane, out.width="768px", out.height="364px"}
knitr::include_graphics('../figures/trad_cors.pdf')
```

Figure 1. Methane oxidation rate is not correlated with functional gene
abundance or ASV richness. Correlations between community attributes and
ecosystem function. A) Abundance of the functional gene pmoA (n = 42) and B)
ASV richness (n = 44). Lines represent the ordinary least squares regression
line with standard errors.

```{r structure, out.width="960px", out.height="1152px"}
knitr::include_graphics('../figures/pcoa_multi.pdf')
```

Figure 2. Microbial community composition is spatially and environmentally
structured. Principle coordinate plots of Bray-Curtis distance representing the
first three axes of community composition. In A, B, and C, points are
identified by Site ID and in D, E, and F, points are identified by wetland or
upland ecosystem. All four environmental covariates separate strongly by
wetland/upland. Axis length is proportional to variance explained as indicated
in parentheses. PC = principal coordinate.

```{r taxa, out.width="768px", out.height="384px"}
knitr::include_graphics('../figures/effect_sizes.pdf')
```

Figure 3. Taxa associated with methane oxidation rates after controlling for
geographic location, environmental similarity, and community composition.
Points are estimates for the linear relationship between the relative abundance
of a single ASV and methane oxidation rate with standard errors from variance
component models including similarity matrices as covariates for community and
environment and site ID for geographic location. Amplicon sequence variants are
labelled at the finest resolution available: genus for all except the Group 1
Acidobacterium. Points are identified by Phylum. Significant taxa were
determined by controlling the false discovery rate at q-value < 0.05.